---
title: "Exercise 3"
author: "Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
library(broom)
library(kableExtra)
library(AER)
library(scales)
library(moderndive)
```

```{r 3data}
ideology <- 
  here::here("data", "CongressRD.csv") %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  rename(obama_2008 = obama2008, 
         gop_2_party_2010 = gop2party2010,
         gop_win_2010 = go_pwin2010)
```

### (a)

### (b)

### (c)

```{r 3c}
ideology %>%
  ggplot(aes(x = gop_2_party_2010, y = ideology)) + 
  geom_point(color = "blue", na.rm = TRUE) + 
  geom_vline(xintercept = .5, color = "red") +
  labs(x = "GOP vote share", y = "Ideology") +
  scale_x_continuous(labels = percent_format(scale = 100))
```


### (d)

$$\operatorname{ideology}_i = \beta_0 + \beta_1 \operatorname{\text{gop_win_2010}}_i + \beta_2 \operatorname{\text{gop_w_party_2010}}_i + \epsilon_i $$
### (e)

```{r 3e}
ideology %>% 
  lm(ideology ~ gop_win_2010 + I(gop_2_party_2010 - .50), .) %>% 
  jtools::summ(digits = 4)

ideology %>% 
  mutate(gop_win_2010 = factor(gop_win_2010, 
                               labels = c("Democrat Win", "Republican Win"))) %>% 
  drop_na() %>% 
  ggplot(aes(x = gop_2_party_2010, y = ideology, color = factor(gop_win_2010))) + 
  geom_point() + 
  geom_parallel_slopes(se = FALSE) +
  labs(x = "GOP vote share", y = "Ideology") +
  scale_x_continuous(labels = percent_format(scale = 100)) +
  theme(legend.title = element_blank(), 
        legend.position = c(.89, .12),
        legend.background = element_blank())
```


### (f)

```{r 3f}
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .50), .) %>% 
  jtools::summ(digits = 4)

ideology %>% 
  mutate(gop_win_2010 = factor(gop_win_2010, 
                               labels = c("Democrat Win", "Republican Win"))) %>% 
  drop_na() %>% 
  ggplot(aes(x = gop_2_party_2010 - .5, y = ideology, color = factor(gop_win_2010))) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "GOP vote share", y = "Ideology") +
  scale_x_continuous(labels = percent_format(scale = 100)) +
  theme(legend.title = element_blank(), 
        legend.position = c(.89, .12),
        legend.background = element_blank()) +
  annotate(geom = "text", x = -.5, y = -.578, label = "A", color = "blue") + 
  annotate(geom = "text", x = 0, y = -.313, label = "B", color = "blue") +
  annotate(geom = "text", x = 0, y = .668, label = "C", color = "red") + 
  annotate(geom = "text", x = .5, y = .686, label = "D", color = "red")
```  

```{r 3f-2, echo=FALSE}
beta <- 
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .50), .) %>% 
  broom::tidy() %>% 
  pull(estimate)
```

$\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1T_i + \hat\beta_2(X_{1i} - C) + \hat\beta_3(X_{1i} - C)T_i$

Consider the following four individuals

1. A democrat with 100% of the vote: point A on the graph above. 
    + $T_i = 0$ 
    + $(X_{1i} - C) = -.5$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_2(X_{1i} - C)$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[3]*(-.5)`$
2. A democrat with 50% of the vote: point B on the graph above.
    + $T_i = 0$ 
    + $(X_{1i} - C) = 0$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1]`$
3. A republican with 50% of the vote: point C on the graph above.
    + $T_i = 1$ 
    + $(X_{1i} - C) = 0$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[2]`$
4. A republican with 100% of the vote: point D on the graph above.
    + $T_i = 1$ 
    + $(X_{1i} - C) = .5$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1T_i + \hat\beta_2(X_{1i} - C) + \hat\beta_3(X_{1i} - C)T_i$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[2] + beta[3]*.5 + beta[4]*.5`$

### (g)

```{r 3g}
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010), .) %>% 
  jtools::summ(digits = 4)

ideology %>% 
  mutate(gop_win_2010 = factor(gop_win_2010, 
                               labels = c("Democrat Win", "Republican Win"))) %>% 
  drop_na() %>% 
  ggplot(aes(x = gop_2_party_2010, y = ideology, color = factor(gop_win_2010))) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "GOP vote share", y = "Ideology") +
  scale_x_continuous(labels = percent_format(scale = 100)) +
  theme(legend.title = element_blank(), 
        legend.position = c(.89, .12),
        legend.background = element_blank())
```

```{r 3g-2}
beta <- 
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010), .) %>% 
  broom::tidy() %>% 
  pull(estimate)

```

$\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1T_i + \hat\beta_2X_{1i} + \hat\beta_3X_{1i}T_i$

Consider the following four individuals

1. A democrat with 100% of the vote:
    + $T_i = 0$ 
    + $X_{1i} = 0$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1]`$
2. A democrat with 50% of the vote: 
    + $T_i = 0$ 
    + $X_{1i} = .50$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat{\beta_2}*.5$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[3]*.5`$
3. A republican with 50% of the vote: point C on the graph above.
    + $T_i = 1$ 
    + $X_{1i} = .5$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1 + \hat\beta_2*.5 + \hat\beta_3*.5$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[2] + beta[3]*.5 + beta[4]*.5`$
4. A repubican with 100% of the vote: point D on the graph above.
    + $T_i = 1$ 
    + $X_{1i} = 1$
    + $\widehat{\operatorname{ideology}_i} = \hat\beta_0 + \hat\beta_1 + \hat\beta_2 + \hat\beta_3$
    + $\widehat{\operatorname{ideology}_i} = `r beta[1] + beta[2] + beta[3]*1 + beta[4]*1`$

To calculate the difference after at the cutoff: 

The jump at the cutoff is distance from #2 to #3 above. So, $\hat\beta_0 + \hat\beta_1 + \hat\beta_2*.5 + \hat\beta_3*.5 - (\hat\beta_0 + \hat{\beta_2}*.5) = \hat\beta_1+\hat\beta_3*.5 = `r beta[2] + beta[4]*.5`$

### (h)

```{r 3h}
ideology %>% 
  drop_na() %>% 
  ggplot(aes(x = gop_2_party_2010)) + 
  geom_histogram(bins = 20, fill = "blue", color = "black") + 
  geom_vline(xintercept = .5, color = "red") +
  labs(x = "GOP vote share", y = "") +
  scale_x_continuous(labels = percent_format(scale = 100))
  
```


### (i)

```{r 3i}
ideology %>% 
  lm(cbind(child_poverty, median_income, obama_2008, white_pct) ~ 
       gop_win_2010 + I(gop_2_party_2010 - .5), .) %>% 
  broom::tidy() %>% 
  filter(term == "gop_win_2010") %>% 
  select(response, statistic, p.value)
```

#### (j)

```{r 3j}
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .50) + 
       child_poverty + median_income + obama_2008 + white_pct, .) %>% 
  jtools::summ(digits = 4)
```

#### (k)

```{r 3k}
ideology %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .50) + 
       gop_win_2010 * I((gop_2_party_2010-.5)^2) + 
       child_poverty + median_income + obama_2008 + white_pct, .) %>% 
  jtools::summ(digits = 4)
```

#### (l)

```{r 3l}
jtools::export_summs(
  ideology %>% 
  filter(between(gop_2_party_2010, .4, .6)) %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .5), .),
  
  ideology %>% 
  filter(between(gop_2_party_2010, .4, .6)) %>% 
  lm(ideology ~ gop_win_2010 * I(gop_2_party_2010 - .5) + 
       child_poverty + median_income + obama_2008 + white_pct, .),
  digits = 4
)

```


### (m)